---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
# Load packages
library(ggplot2)
library(rethinking)
library(cowplot)

# Define grid
dens=100
p_grid = seq(from=0, to =1, length.out = dens)

# Define prior
uniform_prior = rep(1,dens)

# Compute likelihood
likelihood = dbinom(3, size = 6, prob = p_grid)

# Compute posterior 
unstd.posterior = likelihood*uniform_prior

# Stadardize
posterior = unstd.posterior/sum(unstd.posterior)

# Draw plot
Data=data.frame(grid=p_grid,posterior=posterior,prior=uniform_prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

# Performing the analysis with quadratic estimation 

Riccardo.qa = map(alist(w ~ dbinom(6,p), p ~ dunif(0,1)), data=list(w=3)) # Performing quadratic estimation

precis(Riccardo.qa) # getting summary

# Visualizing quadratic estimation in comparrison to analytic solution
c <- 3
q <- 6
curve( dbeta( x , c+1 , q-c+1 ) , from=0 , to=1 )
curve(dnorm( x, 0.5, 0.2), lty=2, add = TRUE)


# To get samples
samples = sample(p_grid, prob = Data$posterior, size = 1e4, replace = TRUE)

# To plot samples
plot(samples)
dens(samples)

# To get the probability < 0.5 from the samples
sum(Data$posterior[ p_grid < 0.5 ]) 

#There is a 50 % percent chance that Riccardo's knowledge is above chance level 
```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
# Creating a function to perform grid approximation
grid_function=function(prior, correct, questions){
  # Compute likelihood
  likelihood = dbinom(correct, size = questions, prob = p_grid)

  # Compute posterior 
  unstd.posterior = likelihood*prior

  # Stadardize
  posterior = unstd.posterior/sum(unstd.posterior)
  
  Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
  
  return(Data)
}


Kristian_grid=grid_function(uniform_prior, 2, 2)

ggplot(Kristian_grid,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

Josh_grid =grid_function(uniform_prior, 160, 198)

ggplot(Josh_grid,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

Mikkel_grid=grid_function(uniform_prior, 66, 132)

ggplot(Mikkel_grid,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
# Creating a more optimistic prior
optimistic_prior= dnorm(p_grid, mean = 0.8, sd = 0.2)

# To create posterior for each teacher

Riccardo_opt = grid_function(optimistic_prior, 3, 6)

ggplot(Riccardo_opt,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

Kristian_opt=grid_function(optimistic_prior, 2, 2)

ggplot(Kristian_opt,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

Josh_opt =grid_function(optimistic_prior, 160, 198)

ggplot(Josh_opt,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

Mikkel_opt=grid_function(optimistic_prior, 66, 132)

ggplot(Mikkel_opt,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

```

Riccardo's and Mikkel's posterior distribution are a more optimistic when using the optimistic prior. Kristian's posterior distribution becomes more pessimistic. Josh's posterior seem unaffected by the optimistic prior which makes sense as the prior corresponds roughly to the data.  

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
# Using a uniform prior 
Riccardo_grid100=grid_function(uniform_prior, 300, 600)
Mikkel_grid100=grid_function(uniform_prior, 6600, 13200)
Josh_grid100 =grid_function(uniform_prior, 16000, 19800)
Kristian_grid100=grid_function(uniform_prior, 200, 200)

R_uni_plot=ggplot(Riccardo_grid100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

M_uni_plot=ggplot(Mikkel_grid100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

J_uni_plot=ggplot(Josh_grid100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

K_uni_plot=ggplot(Kristian_grid100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

# Using an optimistic prior
Riccardo_opt100=grid_function(optimistic_prior, 300, 600)
Mikkel_opt100=grid_function(optimistic_prior, 6600, 13200)
Josh_opt100 =grid_function(optimistic_prior, 16000, 19800)
Kristian_opt100=grid_function(optimistic_prior, 200, 200)

K_opt_plot=ggplot(Kristian_opt100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

R_opt_plot=ggplot(Riccardo_opt100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

J_opt_plot=ggplot(Josh_opt100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

M_opt_plot=ggplot(Mikkel_opt100,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

# To compare plots
plot_grid(M_opt_plot, M_uni_plot)
plot_grid(K_opt_plot, K_uni_plot)
plot_grid(J_opt_plot, J_uni_plot)
plot_grid(R_opt_plot, R_uni_plot)

# Prior is less important having more data

```

There is little difference between the results using the uniform and the optimistic prior when adding more data. We can therefor conclude that as we add more data the influence of the prior becomes less. Also adding more data makes the distribution more narrow. 

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

We would use a skeptic prior (fx mean=0.5, sd = 0.2) indicating that we think their answer will be at chance level. 

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



